---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---
BIS 581 - Product Sentiment Analysis



# load libraries that you feel you need and explain why
```{r codechunk-libraryload, echo=FALSE, message=FALSE, warning=FALSE}

if (!("knitr" %in% installed.packages())) install.packages("knitr")

library(knitr)

# For text mining and natural language processing in a tidy data format.
library(tidytext) 

# Helps in reshaping and cleaning the data. 
library(tidyr) 

# Creating elegant network visualizations.
library(ggraph) 

# For graph/network analysis and visualization.
library(igraph) 

# For generating a word cloud to visualize word frequencies.
library(wordcloud)

# Provides a grammar of data manipulation verbs such as filter, select, and mutate.
library(dplyr)

# Provides text datasets for natural language processing, such as sentiment lexicons and pre-trained word embeddings.
library(textdata)
library(RColorBrewer)
library(wordcloud)
# Install and load stopwords package if not installed
if (!requireNamespace("stopwords", quietly = TRUE)) install.packages("stopwords")

# Load the package
library(stopwords)
# Install and load the required package
if (!requireNamespace("stringr", quietly = TRUE)) install.packages("stringr")

# Load stringr package
library(stringr)

# Install and load the required package
if (!requireNamespace("tm", quietly = TRUE)) install.packages("tm")

# Load tm package
library(tm)


```

# Load the data from a file
```{r code1, echo=FALSE, message=FALSE, warning=FALSE}
options(stringsAsFactors = FALSE)

#place your code file and data file in the same folder OR use this code and adjust it for your environment
#folderst <- "D://data//myproject//"
#setwd(folderst)

posts <- readRDS("productsalescomments.rds")

```


# Assess the data
```{r}
# Viewing the dataset
View(posts)

```

```{r}
# Viewing the structure of dataset
str(posts)

# Display the first few rows
head(posts)
```
# Is any data prep needed?   If so, adjust data here and comment on why in your submission
```{r}
# Cleaning T_Price and T_Cost columns by removing $ symbols and converting to numeric
# To enable numerical analysis, as price and cost are currently stored as characters.
posts_clean <- posts %>%
  mutate(
    # Removing non-numeric characters for calculations.
    T_Price_numeric = as.numeric(gsub("[^0-9.]", "", T_Price)), 
    T_Cost_numeric = as.numeric(gsub("[^0-9.]", "", T_Cost)),
    
    # Converting back to character format with '$'.
    T_Price = paste0("$", format(T_Price_numeric, nsmall = 2)), 
    T_Cost = paste0("$", format(T_Cost_numeric, nsmall = 2))  
  )


# Filling missing values in numeric columns with 0. To avoid issues with calculations due to missing values and maintain data consistency.
numeric_columns <- c("Star_Rating", "T_Price_numeric", "T_Cost_numeric")
posts_clean <- posts_clean %>%
  mutate(across(all_of(numeric_columns), ~ ifelse(is.na(.), 0, .)))

# Filling missing values in the Comment column with 'Unknown' and convert text to lowercase
posts_clean <- posts_clean %>%
  mutate(
    # Replacing missing and blank comments
    Comment = ifelse(is.na(Comment) | Comment == "", "unknown", Comment),
    
    # Replacing missing CustomerID with 0
    CustomerID = ifelse(is.na(CustomerID), 0, CustomerID),
    
    # Converting all character fields to lowercase
    across(where(is.character), tolower)                                    
  )

# Removing duplicate rows if any exist. To ensure each transaction is only counted once for accurate analysis.
posts_clean <- posts_clean %>% distinct()

str(posts_clean)
head(posts_clean)

```

```{r}
View(posts_clean)
```

# Decide if you want to filter
```{r}

# Filter out rows without meaningful comments or customer information
# Filtering out rows where Comment is 'unknown' to focus on meaningful feedback.
# Filtering out rows where CustomerID is '0' to ensure we can analyze customer behavior.
posts_filtered <- posts_clean %>%
  filter(Comment != "unknown", CustomerID != 0)
str(posts_filtered)
head(posts_filtered)

```
```{r}
# Define a list of common ice cream-related words and phrases to remove
remove_words <- c("ice", "cream", "vanilla", "chocolate", "strawberry", 
                  "flavor", "cone", "scoop", "dessert", "topping", 
                  "milkshake", "gelato", "sundae", "frozen", "custard", 
                  "i'm", "too")  # Added "i'm" and "too"

# 🔥 Step 1: Tokenize Comments and Remove Unwanted Words
cleaned_comments <- posts_filtered %>%
  unnest_tokens(word, Comment) %>%
  filter(!word %in% remove_words) %>%
  group_by(CustomerID) %>%
  summarize(Clean_Comment = paste(word, collapse = " "), .groups = "drop")

# 🔥 Step 2: Merge Cleaned Comments Back Into Original Dataset
posts_filtered_cleaned <- posts_filtered %>%
  select(-Comment) %>%  # Remove original Comment column
  left_join(cleaned_comments, by = "CustomerID") %>%  # Merge cleaned comments
  rename(Comment = Clean_Comment)  # Rename back to Comment

# 🔥 Step 3: Ensure All Columns Are Present
head(posts_filtered_cleaned)


```




# Now answer/create the following:

1. Top 15 meaningful words
```{r}
# Create a custom stopword list
custom_stopwords <- c(stopwords("en"), "i'm", "im", "m", "s", "t", "ve", "re", "ll", "d", "just", "like")  # Removes unwanted short words

# 🔥 Step 1: Pre-clean the text before tokenization
posts_filtered_cleaned <- posts_filtered_cleaned %>%
  mutate(Comment = tolower(Comment),  # Convert everything to lowercase
         Comment = str_replace_all(Comment, "\\b(i'm|im)\\b", ""),  # Remove "i'm" and "im"
         Comment = removeWords(Comment, custom_stopwords),  # Remove all stopwords
         Comment = str_replace_all(Comment, "[[:punct:]]", ""),  # Remove punctuation
         Comment = str_squish(Comment))  # Remove extra spaces

# 🔥 Step 2: Tokenization and Further Cleanup
top_words <- posts_filtered_cleaned %>%
  unnest_tokens(word, Comment) %>%
  filter(nchar(word) > 1) %>%  # Removes single-letter words except "a" and "i"
  count(word, sort = TRUE) %>%
  slice_max(n, n = 15)

# ✅ Step 3: Debugging Check
print(unique(top_words$word))  # 🔍 Verify if "m" and "s" are removed

# ✅ Step 4: View Cleaned Top 15 Words
top_words
```

2. Word cloud
```{r}
set.seed(123)

# Tokenize comments to words, remove stop words, and count word frequencies
tokenized_comments <- posts_filtered_cleaned %>%
  unnest_tokens(word, Comment) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)

# Generating the word cloud with a custom color palette
wordcloud(words = tokenized_comments$word, freq = tokenized_comments$n, min.freq = 5,
          max.words = 100, random.order = FALSE, scale = c(3, 0.5), 
          colors = brewer.pal(8, "Dark2"))

```


3. Who are the top 5 customers by CustomerID who posts the most comments on products?
```{r}
# Finding the top 5 customers by number of comments with tiebreaker by CustomerID
top_customers <- posts_filtered_cleaned %>%
  count(CustomerID, sort = TRUE) %>%
  arrange(desc(n), CustomerID) %>% # Sorting by n, then by CustomerID for tiebreaking
  slice_head(n = 5)                # Select the top 5 customers only

top_customers

```


4. For each these 5 are these positive or negative customers overall considering a measure using the afinn sentiment measure?
```{r}
afinn <- get_sentiments("afinn")

# Calculating AFINN Sentiment Score for Top 5 Customers
sentiment_top_customers <- posts_filtered_cleaned %>%
  filter(CustomerID %in% top_customers$CustomerID) %>%
  unnest_tokens(word, Comment) %>%
  inner_join(afinn, by = "word") %>%
  group_by(CustomerID) %>%
  summarize(sentiment_score = sum(value)) %>%
  arrange(desc(sentiment_score))

# Calculating Average Star Rating for Top 5 Customers
average_rating_top_customers <- posts_filtered_cleaned %>%
  filter(CustomerID %in% top_customers$CustomerID) %>%
  group_by(CustomerID) %>%
  summarize(avg_rating = mean(Star_Rating, na.rm = TRUE))

# Combining Sentiment Scores and Ratings, and Classify Sentiment
combined_sentiment <- sentiment_top_customers %>%
  inner_join(average_rating_top_customers, by = "CustomerID") %>%
  mutate(overall_sentiment = case_when(
    avg_rating < 3 ~ "negative",
    avg_rating >= 3 ~ "positive"
  ))

# View the overall sentiment for top customers
combined_sentiment

```

5. What Parlor site has the most comments?  Are these negative or positive?

```{r}
# Finding the Parlor Site with the Most Comments
most_comments_parlor <- posts_filtered_cleaned %>%
  count(ParlorLocation, sort = TRUE) %>%
  slice_max(n, n = 1)

# Viewing the Parlor Site with the Most Comments
most_comments_parlor

# Filtering for the Parlor Site with the Most Comments
parlor_with_most_comments <- posts_filtered_cleaned %>%
  filter(ParlorLocation == most_comments_parlor$ParlorLocation)

# Calculating Sentiment Score Using AFINN
parlor_sentiment <- parlor_with_most_comments %>%
  unnest_tokens(word, Comment) %>%
  inner_join(afinn, by = "word") %>%
  summarize(sentiment_score = sum(value))

# Calculating Average Rating for the Parlor Site
avg_rating_parlor <- parlor_with_most_comments %>%
  summarize(avg_rating = mean(Star_Rating, na.rm = TRUE))

# Determining Overall Sentiment for the Parlor Site
overall_sentiment <- parlor_sentiment %>%
  mutate(
    ParlorLocation = most_comments_parlor$ParlorLocation,
    num_comments = most_comments_parlor$n,
    avg_rating = avg_rating_parlor$avg_rating,
    overall_sentiment = case_when(
      avg_rating < 3 ~ "negative",
      avg_rating >= 3 ~ "positive"
    )
  )

# View the Overall Sentiment for the Parlor with the Most Comments
overall_sentiment


```

6. What production site has the most comments?  Are these negative or positive?
```{r}

# Finding the Production Site with the Most Comments
most_comments_production <- posts_filtered_cleaned %>%
  count(ProductionLocation, sort = TRUE) %>%
  slice_max(n, n = 1)

# View the Production Site with the Most Comments
most_comments_production

# Filtering for the Production Site with the Most Comments
production_with_most_comments <- posts_filtered_cleaned %>%
  filter(ProductionLocation == most_comments_production$ProductionLocation)

# Calculating Sentiment Score Using AFINN
production_sentiment <- production_with_most_comments %>%
  unnest_tokens(word, Comment) %>%
  inner_join(afinn, by = "word") %>%
  summarize(sentiment_score = sum(value))

# Calculating Average Rating for the Production Site
avg_rating_production <- production_with_most_comments %>%
  summarize(avg_rating = mean(Star_Rating, na.rm = TRUE))

# Determining Overall Sentiment for the Production Site
overall_production_sentiment <- production_sentiment %>%
  mutate(overall_sentiment = case_when(
    avg_rating_production$avg_rating < 3 ~ "negative",
    avg_rating_production$avg_rating >= 3 ~ "positive"
  ))

# View the Overall Sentiment for the Production Site with the Most Comments
overall_production_sentiment

```

7. What are the top 10 most frequent comments made by customers and how many times did each comment get made?
```{r}
# Step 1: Count the Frequency of Each Comment
top_comments <- posts_filtered_cleaned %>%
  group_by(Comment) %>%
  summarize(comment_count = n()) %>%
  arrange(desc(comment_count))

# Step 2: Select the Top 10 Most Frequent Comments Without Ties
top_10_comments <- top_comments %>%
  slice_max(comment_count, n = 10, with_ties = FALSE)

# View the Top 10 Most Frequent Comments
top_10_comments


```

8. Create a network diagram based on bigrams
```{r}
# Tokenize Comments into Bigrams and Remove Stop Words
bigrams <- posts_filtered_cleaned %>%
  unnest_tokens(bigram, Comment, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word, !word2 %in% stop_words$word) %>%
  filter(word1 != "NA" & word2 != "NA") # Remove any "NA" values

# Count the Frequency of Each Bigram
bigram_counts <- bigrams %>%
  count(word1, word2, sort = TRUE) %>%
  filter(n > 3) # Lower frequency threshold to only include more frequent bigrams

# Limit to Top 15 Most Frequent Bigrams
bigram_top_15 <- bigram_counts %>%
  slice_max(n, n = 15)

# Create a Graph from the Bigrams
bigram_graph <- bigram_top_15 %>%
  graph_from_data_frame()

# Plot the Network Diagram
set.seed(123)
ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, edge_width = 1.5) +
  geom_node_point(color = "steelblue", size = 8) + # Simplify node color and make nodes larger
  geom_node_text(aes(label = name), repel = TRUE, vjust = 1, hjust = 1, size = 5, color = "black") + # Use larger labels for clarity
  theme_void() +
  labs(title = "Network Diagram")
```


GRADS----
assuming you work for the company/organization for which these product comments have been collected, 
what can you infer from the data? 
If the company was asking you if they should take any actions based on customer feedback, 
what would you tell them and why?

Actions to be taken:
Marketing Campaigns: Focus on popular flavors to boost interest and attract new customers.

Customer Testimonials: Use positive customer feedback in promotional material.

Supply Chain Improvements: Ensure the availability of popular products to meet demand.
Highlight and promote seasonal or uncommon flavors to generate excitement.
